---
title: "Development of MASLD Treatment Model Assumptions"
bibliography: ../../references/model_assumptions_references.bib
author: "<a href='https://github.com/ASW-Analyst/'>Andy Wilson</a><br><a href='https://transformationunit.nhs.uk/'>NHS Transformation Unit</a>"
date: "01/15/2025"
date-format: "DD MMMM YYYY"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    self-contained: true
    css: ../../../src/config/nhse_theme.css
    number-sections: true

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = '100%'
)

library(here)

```

```{r logo}

htmltools::img(src = knitr::image_uri(paste0(here(), "/images/nhse_logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:5%; padding:10px;',
               width = "180px",
               heigth = "180px")

```

# Introduction

::: {.callout-note}
This is a live document that will be continually updated during the course of this project. Literature review is currently on-going to establish assumptions and sensitivies to be applied within the model.
:::

The purpose of this document is to identify the relevant and appropriate sources for each of the MASLD treatment model assumptions. This will include:

* An explanation of __why__ this assumption needs to be determined.
* Discussion of literature that provides estimates of the assumption including the __geographical area__, __time period__ of the data collection and any important __limitations__.
* The __suggested value__ of this assumption to be used in the model and the range of __sensitivities__ to be applied.


***
<br/>

# Starting Population Assumptions Development
This section discusses the development of assumptions related to determining the initial starting populations of eligible patients. The flow of determining the starting population are shown in the diagram below:

![Starting Population Generation](../diagrams/masld_treatment_modelling_plan.drawio.svg)

## Prevalence and Sub-Population Prevalence

### Why is this required?
The prevalence of MASLD & MASH is necessary to establish the population of patients that could be considered for each of the four treatment options. By establishing the prevalence this will enable the initial population to be calculated for MASLD patients across England. Further assumptions will then be applied to determine the percentage of this population that is eligible for the treatment.

### Findings from Literature

The table below provides an overview of the estimates of the prevalence of MASLD and MASH from relevant literature:

| **Reference**                 | **Estimate of Prevalence (95%CI)**| **Location**  | **Comments**                        |
|-------------------------------|-----------------------------------|---------------|-------------------------------------|
| [@lekakis2024naturalhistmasld]| MASLD: 32% (30 - 35%)             | Global        | Meta-analysis of NAFLD worldwide    |
| [@younossi2016globalnafld]    | MASLD: 25.2% (22.1 - 28.7%)       | Global        | Meta-analysis of NAFLD              |
| [@younossi2019globalhealth]   | MASLD: 25%                        | Global        | Review of global risk factors       |
| [@younossi2019globalhealth]   | MASLD: 24%                        | Europe        | Review of global risk factors       |
| [@younossi2023systematic]     | MASLD: 25.1% (20.6 - 30.3%)       | Western Europe| Systematic Review                   |
| [@younossi2023systematic]     | MASLD: 38.0% (33.7 - 42.5%)       | Global        | Systematic Review - 2016 to 2019    |
| [@alenezi2022ukprev]          | MASLD: 23.7% (20.6 - 26.8%)       | UK            | Systematic Review and meta-analysis |
| [@younossi2023systematic]     | MASH: 5.27%                       | Global        | Systematic Review                   |
| [@younossi2023systematic]     | MASH: 4.02%                       | Western Europe| Systematic Review                   |
| [@younossi2023systematic]     | MASH: 4.02%                       | Western Europe| Systematic Review                   |
| [@younossi2019globalhealth]   | MASH: 3 - 5%                      | Global        | Review of global risk factors       |


### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Fibrosis Stage

### Why is this required?
This is required to identify the relevant populations of patients with MASLD and MASH for each Fibrosis stage. It is anticipated that patients will be eligible for specific treatments based on their Fibrosis Stage.

### Findings from Literature
Findings from relevant literature are shown in the table below:

| **Reference**                 | **Sub-group** | **Estimate (95% CI) **| **Location**  | **Comments**                    |
|-------------------------------|---------------|-----------------------|---------------|---------------------------------|
| [@basu2022mgmt]               | MASH          | 20% (of MAFLD)        | USA           | Review of management of NAFLD   |
| [@basu2022mgmt]               | F3/F4         | 20% (of MASH)         | USA           | Review of management of NAFLD   |
| [@singh2015progression]       | F0            | 35.8%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F1            | 32.5%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F2            | 16.7%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F3            | 9.3%                  | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F4            | 5.7%                  | Global        | Biopsy proven NAFLD             |

### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Co-morbidities and Risk Factors

### Why is this required?
It is important to consider the expected overlap of patients with MASLD and other conditions such as Obesity or Diabetes. Patients who have MASLD but are being treated on an obesity pathway with a GLP-1 agonist may not be prescribed one of the four treatments in addition to their obesity treatment. Therefore, an estimate of patients with MASLD to exclude from the treatment population is required.

### Findings from Literature

#### Obesity

The table below shows the estimated prevalence of **obesity for patients diagnosed with MASLD or MASH**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@basu2022mgmt]         | MASLD             | 51%                               | USA           | MASLD patients with obesity |
| [@pustjens2024profile]  | MASLD - All ages  | 89.5%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 20-40     | 83.1%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 40-60     | 92.4%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 60-80     | 92.9%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@basu2022mgmt]         | MASH              | 82%                               | USA           | MASH patients with obesity  |

#### Type 2 Diabetes
The table below shows the estimated prevalence of **Type 2 Diabetes for patients with MASLD or MASH**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@konyn2022mortality]   | MASLD             | 22.5% (17.9 - 27.9%)              | USA           | MASLD patients with T2D     |
| [@konyn2022mortality]   | MASH              | 43.6% (30.3 - 58.0%)              | USA           | MASH patients with T2D      |
| [@pustjens2024profile]  | MASLD - All ages  | 66.8%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 20-40     | 37.4%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 40-60     | 74.9%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 60-80     | 88.1%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |

The table below shows the estimated prevalence of **MASLD, MASH or Fibrosis Stage for patients with Type 2 Diabetes**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@basu2022mgmt]         | MASLD             | 60%                               | USA           | T2D patients with MASLD     |
| [@cho2023diabetes]      | MASLD             | 65.04% (61.79 - 68.15%)           | Global        | T2D patients with MASLD     |
| [@konyn2022mortality]   | MASLD             | 55.5% (57.3 - 63.7%)              | USA           | T2D patients with MASLD     |
| [@cho2023diabetes]      | MASH              | 31.55% (17.12 - 50.70%)           | Global        | T2D patients with MASH      |
| [@konyn2022mortality]   | MASH              | 37.3% (24.7 - 50.0%)              | USA           | T2D patients with MASH      |
| [@cho2023diabetes]      | F2-F4             | 35.54% (19.56 - 55.56%)           | Global        | T2D patients with MASLD F2-F4|
| [@cho2023diabetes]      | F3-F4             | 14.95% (11.03 - 19.95%)           | Global        | T2D patients with MASLD F3-F4|

### Suggested Model Inputs and Sensitivites

**TBC**


***
<br/>

# Temporal Assumptions Development
This section discusses the development of assumptions related to temporal elements of the model. The flow of determining appropriate assumptions are described in the diagram below:

![Temporal Assumptions](../diagrams/masld_treatment_modelling_plan-Temporal Requirements.drawio.svg)

## MASLD Incidence

### Why is this required?
The modelling of future treatments will need to run over a number of years, this modelling period is still to be confirmed. Therefore, consideration needs to be given to the number of new patients diagnosed with MASLD who may become eligible for treatment each year, i.e. the incidence rate. This will ensure that modelling future year treatment activities and associated costs account accurately for future demand.

### Findings from Literature

### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## MASLD and MASH Recovery and Progression

### Why is this required?
The modelling of future treatments will need to run over a number of years, this modelling period is still to be confirmed. Therefore, consideration needs to be given to how patients in each of the MASLD and MASH sub-groups will progress and recover during this period.

For example, _how many patients with a Fibrosis score of 2 might progress to a Fibrosis score of 3 during that year and no-longer be eligible for a specific treatment?_ Likewise consideration will need to be given to patient's who see an improvement in the condition of their liver through new treatments or through lifestyle changes.

### Findings from Literature

The table below provides an overview of the proportion of patients who see disease progression including the progression of Fibrosis:

| **Reference**           | **Group**             | **Estimate (95% CI)** | **Location**     | **Comments**                  |
|-------------------------|-----------------------|-----------------------|------------------|-------------------------------|
| [@basu2022mgmt]         | MASH Patients         | 40%                   | USA              | Fibrosis progression          |
| [@estes2018modelling]   | F0 => F1 Male (<40)   | 0.6% (0.3-0.9%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F0 => F1 Female (<40) | 0.5% (0.3-0.7%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F0 => F1 Male (40+)   | 1.5% (0.9-2.4%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F0 => F1 Female (40+) | 1.3% (0.8-2.0%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F1 => F2 Male (<40)   | 4.5% (2.7-6.9%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F1 => F2 Female (<40) | 3.8% (2.2-5.8%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F1 => F2 Male (40+)   | 11.9% (7.0-18.3%)     | UK               | Annual transition probability |
| [@estes2018modelling]   | F1 => F2 Female (40+) | 10.0% (5.9-15.3%)     | UK               | Annual transition probability |
| [@estes2018modelling]   | F2 => F3 Male (<40)   | 4.5% (2.7-6.9%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F2 => F3 Female (<40) | 3.8% (2.2-5.8%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F2 => F3 Male (40+)   | 11.9% (7.0-18.3%)     | UK               | Annual transition probability |
| [@estes2018modelling]   | F2 => F3 Female (40+) | 10.0% (5.9-15.3%)     | UK               | Annual transition probability |
| [@estes2018modelling]   | F3 => F4 Male (<40)   | 5.5% (3.1-10.4%)      | UK               | Annual transition probability |
| [@estes2018modelling]   | F3 => F4 Female (<40) | 4.6% (2.6-8.7%)       | UK               | Annual transition probability |
| [@estes2018modelling]   | F3 => F4 Male (40+)   | 8.9% (5.1-17.0%)      | UK               | Annual transition probability |
| [@estes2018modelling]   | F3 => F4 Female (40+) | 7.4% (4.2-14.1%)      | UK               | Annual transition probability |

### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Retention

### Why is this required?
Consideration needs to be given to how many patients will be retained and remain receiving treatment each year. Some patients may decline continuing treatment so this needs to be accounted for.

### Findings from Literature

### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Mortality

### Why is this required?
Consideration needs to be given to how many patients will die during each year of the model and are therefore, no longer requiring treatment.

### Findings from Literature

### Suggested Model Inputs and Sensitivites

**TBC**

***
<br/>

# References

::: {#refs}
:::