---
title: "Development of MASLD Treatment Model Assumptions"
bibliography: ../../references/model_assumptions_references.bib
author: "<a href='https://github.com/ASW-Analyst/'>Andy Wilson</a><br><a href='https://transformationunit.nhs.uk/'>NHS Transformation Unit</a>"
date: "02/09/2025"
date-format: "DD MMMM YYYY"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    self-contained: true
    css: ../../../src/config/nhse_theme.css
    number-sections: true

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = '100%'
)

library(here)

```

```{r logo}

htmltools::img(src = knitr::image_uri(paste0(here(), "/images/nhse_logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:5%; padding:10px;',
               width = "180px",
               heigth = "180px")

```

# Introduction

::: {.callout-note}
This is a live document that will be continually updated during the course of this project. Literature review is currently on-going to establish assumptions and sensitivties to be applied within the model.
:::

The purpose of this document is to identify the relevant and appropriate sources for each of the MASLD treatment model assumptions. This will include:

* An explanation of __why__ this assumption needs to be determined.
* Discussion of literature that provides estimates of the assumption including the __geographical area__, __time period__ of the data collection and any important __limitations__.
* The __suggested value__ of this assumption to be used in the model and the range of __sensitivities__ to be applied.


***
<br/>

# Starting Population Assumptions Development
This section discusses the development of assumptions related to determining the initial starting populations of eligible patients. The flow of determining the starting population are shown in the diagram below:

![Starting Population Generation](../diagrams/masld_treatment_modelling_plan.drawio.svg)

## Prevalence and Sub-Population Prevalence

### Why is this required?
The prevalence of MASLD & MASH is necessary to establish the population of patients that could be considered for each of the four treatment options. By establishing the prevalence this will enable the initial population to be calculated for MASLD patients across England. Further assumptions will then be applied to determine the percentage of this population that is eligible for the treatment.

### Findings from Literature

The table below provides an overview of the estimates of the prevalence of MASLD and MASH from relevant literature:

| **Reference**                 | **Estimate of Prevalence (95%CI)**| **Location**  | **Comments**                        |
|-------------------------------|-----------------------------------|---------------|-------------------------------------|
| [@lekakis2024naturalhistmasld]| MASLD: 32% (30 - 35%)             | Global        | Meta-analysis of NAFLD worldwide    |
| [@younossi2016globalnafld]    | MASLD: 25.2% (22.1 - 28.7%)       | Global        | Meta-analysis of NAFLD              |
| [@younossi2019globalhealth]   | MASLD: 25%                        | Global        | Review of global risk factors       |
| [@younossi2019globalhealth]   | MASLD: 24%                        | Europe        | Review of global risk factors       |
| [@younossi2023systematic]     | MASLD: 25.1% (20.6 - 30.3%)       | Western Europe| Systematic Review                   |
| [@younossi2023systematic]     | MASLD: 38.0% (33.7 - 42.5%)       | Global        | Systematic Review - 2016 to 2019    |
| [@alenezi2022ukprev]          | MASLD: 23.7% (20.6 - 26.8%)       | UK            | Systematic Review and meta-analysis |
| [@younossi2023systematic]     | MASH: 5.27%                       | Global        | Systematic Review                   |
| [@younossi2023systematic]     | MASH: 4.02%                       | Western Europe| Systematic Review                   |
| [@younossi2023systematic]     | MASH: 4.02%                       | Western Europe| Systematic Review                   |
| [@younossi2019globalhealth]   | MASH: 3 - 5%                      | Global        | Review of global risk factors       |


### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Fibrosis Stage

### Why is this required?
This is required to identify the relevant populations of patients with MASLD and MASH for each Fibrosis stage. It is anticipated that patients will be eligible for specific treatments based on their Fibrosis Stage.

### Findings from Literature
Findings from relevant literature are shown in the table below:

| **Reference**                 | **Sub-group** | **Estimate (95% CI) **| **Location**  | **Comments**                    |
|-------------------------------|---------------|-----------------------|---------------|---------------------------------|
| [@basu2022mgmt]               | MASH          | 20% (of MAFLD)        | USA           | Review of management of NAFLD   |
| [@basu2022mgmt]               | F3/F4         | 20% (of MASH)         | USA           | Review of management of NAFLD   |
| [@singh2015progression]       | F0            | 35.8%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F1            | 32.5%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F2            | 16.7%                 | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F3            | 9.3%                  | Global        | Biopsy proven NAFLD             |
| [@singh2015progression]       | F4            | 5.7%                  | Global        | Biopsy proven NAFLD             |
| [@owrangi2024cirrhosis]       | F4            | 3.26% (2.47-4.31%)    | Global        | Prevalence within MASLD         |
| [@owrangi2024cirrhosis]       | F4            | 4.13% (2.88-5.87%)    | Western Europe| Prevalence within MASLD         |
| [@economicimpactuk2018]        | F0            | 22.90%                | UK            | Derived from diagnosis estimates |
| [@economicimpactuk2018]        | F1            | 35.88%                | UK            | Derived from diagnosis estimates |
| [@economicimpactuk2018]        | F2            | 20.23%                | UK            | Derived from diagnosis estimates |
| [@economicimpactuk2018]        | F3            | 12.60%                | UK            | Derived from diagnosis estimates |
| [@economicimpactuk2018]        | F4            | 8.39%                 | UK            | Derived from diagnosis estimates |


### Suggested Model Inputs and Sensitivites

**TBC**

<br/>

## Co-morbidities and Risk Factors

### Why is this required?
It is important to consider the expected overlap of patients with MASLD and other conditions such as Obesity or Diabetes. Patients who have MASLD but are being treated on an obesity pathway with a GLP-1 agonist may not be prescribed one of the four treatments in addition to their obesity treatment. Therefore, an estimate of patients with MASLD to exclude from the treatment population is required.

### Findings from Literature

#### Obesity

The table below shows the estimated prevalence of **obesity for patients diagnosed with MASLD or MASH**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@basu2022mgmt]         | MASLD             | 51%                               | USA           | MASLD patients with obesity |
| [@pustjens2024profile]  | MASLD - All ages  | 89.5%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 20-40     | 83.1%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 40-60     | 92.4%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 60-80     | 92.9%                             | USA           | MASLD patients with obesity (US Abdo confirmed) |
| [@basu2022mgmt]         | MASH              | 82%                               | USA           | MASH patients with obesity  |

#### Type 2 Diabetes
The table below shows the estimated prevalence of **Type 2 Diabetes for patients with MASLD or MASH**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@konyn2023mortality]   | MASLD             | 22.5% (17.9 - 27.9%)              | USA           | MASLD patients with T2D     |
| [@sripongpun2024mortality] | MASLD          | 43.3%                             | USA           | MASH patients with T2D      |
| [@konyn2023mortality]   | MASH              | 43.6% (30.3 - 58.0%)              | USA           | MASH patients with T2D      |
| [@pustjens2024profile]  | MASLD - All ages  | 66.8%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 20-40     | 37.4%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 40-60     | 74.9%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |
| [@pustjens2024profile]  | MASLD - 60-80     | 88.1%                             | USA           | MASLD patients with pre-diabetes (US Abdo confirmed) |

The table below shows the estimated prevalence of **MASLD, MASH or Fibrosis Stage for patients with Type 2 Diabetes**:

| **Reference**           | **Group**         | **Estimated Prevalence (95% CI)** | **Location**  | **Comments**                |
|-------------------------|-------------------|-----------------------------------|---------------|-----------------------------|
| [@basu2022mgmt]         | MASLD             | 60%                               | USA           | T2D patients with MASLD     |
| [@cho2023diabetes]      | MASLD             | 65.04% (61.79 - 68.15%)           | Global        | T2D patients with MASLD     |
| [@konyn2023mortality]   | MASLD             | 55.5% (57.3 - 63.7%)              | USA           | T2D patients with MASLD     |
| [@cho2023diabetes]      | MASH              | 31.55% (17.12 - 50.70%)           | Global        | T2D patients with MASH      |
| [@konyn2023mortality]   | MASH              | 37.3% (24.7 - 50.0%)              | USA           | T2D patients with MASH      |
| [@cho2023diabetes]      | F2-F4             | 35.54% (19.56 - 55.56%)           | Global        | T2D patients with MASLD F2-F4|
| [@cho2023diabetes]      | F3-F4             | 14.95% (11.03 - 19.95%)           | Global        | T2D patients with MASLD F3-F4|

### Suggested Model Inputs and Sensitivites

**TBC**


***
<br/>


## Diagnosis Rates

### Why is this required?
This is required to determine the known population within each sub-group that are currently diagnosed and could be commence treatment.

### Findings from Literature
The table below provides an overview of estimated diagnosis rates for different subgroups:

| **Reference**           | **Group**         | **Estimated Diagnosis Rate (95% CI)**| **Location**  | **Comments**  |
|-------------------------|-------------------|--------------------------------------|---------------|---------------|
| [@economicimpactuk2018] | F0                |2.0%                                  | UK            | Cross-sectional feasibility study |
| [@economicimpactuk2018] | F1                |2.0%                                  | UK            | Cross-sectional feasibility study |
| [@economicimpactuk2018] | F2                |16.5%                                 | UK            | Cross-sectional feasibility study |
| [@economicimpactuk2018] | F3                |58.3%                                 | UK            | Cross-sectional feasibility study |
| [@economicimpactuk2018] | F4                |100%                                  | UK            | Cross-sectional feasibility study |

Additionally, [@Schattenburg2021] determined the diagnosis rates relative to the overall adult population in the UK. These are shown in the table below:

| **Reference**           | **Group**         | **Estimated Diagnosis Rate (95% CI)**| **Location**  | **Comments**  |
|-------------------------|-------------------|--------------------------------------|---------------|---------------|
| [@Schattenburg2021]     | F0                |0.02%                                 | UK            | Relative to adult UK - Higher Prevalence Scenario|
| [@Schattenburg2021]     | F1                |0.03%                                 | UK            | Relative to adult UK - Higher Prevalence Scenario|
| [@Schattenburg2021]     | F2                |0.14%                                 | UK            | Relative to adult UK - Higher Prevalence Scenario|
| [@Schattenburg2021]     | F3                |0.30%                                 | UK            | Relative to adult UK - Higher Prevalence Scenario|
| [@Schattenburg2021]     | F4                |0.31%                                 | UK            | Relative to adult UK - Higher Prevalence Scenario|
| [@Schattenburg2021]     | F0                |0.01%                                 | UK            | Relative to adult UK - Lower Prevalence Scenario|
| [@Schattenburg2021]     | F1                |0.02%                                 | UK            | Relative to adult UK - Lower Prevalence Scenario|
| [@Schattenburg2021]     | F2                |0.07%                                 | UK            | Relative to adult UK - Lower Prevalence Scenario|
| [@Schattenburg2021]     | F3                |0.16%                                 | UK            | Relative to adult UK - Lower Prevalence Scenario|
| [@Schattenburg2021]     | F4                |0.16%                                 | UK            | Relative to adult UK - Lower Prevalence Scenario|

### Suggested Model Inputs and Sensitivites

**TBC**


***
<br/>


# References

::: {#refs}
:::